Centrality and Activity for Envolving Networks

### 2013-06-24 network-science/paper.tex

\definition $a_{\mu,\eta}(G_t,v)$ is the \textbf{evolving activity} for node $v$, if
\begin{equation}
a(G_t,v)=\begin{cases}
0& (z\geq 0)\\
\mu^{1/z} & (z<0)
\end{cases},
\end{equation}
where $0<\mu<1$ is the \textbf{decay rate}, and $z$ is a status variable that can
be determined by the following routine:
\begin{itemize}
\item $z=0$ initially, and $z$ can never exceed $0$.
\item $z$ decrease by $\left(\eta\sum_i w_i\right)$, when new edges with weight
$w_i$'s are created by $v$, as discussed in \eqref{eq-sym}, Here $\eta>0$ is the
\textbf{activity delay}.
\item $z$ increase consistently by time, at the same rate of $t$.
\end{itemize}

@@

Since $0<\mu<1$, and $z\in\mathbb{R}$, we have $0<a(G_t,v)<1$.
\end{proof}

@@

\section{Data structure and Algorithms}
\subsection{Maintaining the basic graph structure}
We formally define the abstract data type for the pure growing model, and
create algorithmic functions to the data structure.

A evolving network $G_t(V,E_t)$ is stored in following contents:
\begin{itemize}
\item \textit{currentTime}, a non-negative real number.
\item $\mu$, a real number in range $(0,1)$.
\item $\eta$, a non-negative real number.
\item \textit{edgeList}[$1\dots|V|$], an array that every of
whose entry points to a sorted array again, consisting the description of the
in edges of $v_i$ in the graph, each with elements:
\begin{itemize}
\item $u$, the other end point of the edge.
\item $t$, the time when this edge is added. (the array is sorted by $t$)
\item $w$, the weight of the edge, can be reduced if the network is
unweighted.
\item $s$, the sum of all previous $w$, including this one.
\item $z$, the status variable for calculating the evolving activity of node
$v_i$ at time $t$ after updating this edge.
\end{itemize}

@@

returns whether $u$ and $v$ is connected in the network at time $t$.
\item \textsc{Query Connectivity}($t,u,v$)

returns whether $u$ and $v$ is connected in the network at time $t$.
\item \textsc{Query Activity}($t,v$)

returns the evolving activity $a_{\mu,\eta}(G_t,v)$.
\item \textsc{Query Centrality}($t$, $\alpha$, $\beta$, $\mathbf{x}_0$, $m$)

returns the active centrality vector $\mathbf{x}(G_t)$. Algorithm
\ref{alg-cen} use the iteration method start with $\mathbf{x}_0$ and for
$m$ times.
\end{itemize}

\begin{algorithm}[htbp]
\caption{\textsc{Add Edge}($v_i,v_j,t,w$)}
\label{alg-add}
\textbf{Time Complexity:} $O(\log |V|)$\\
\textbf{Additional Space:} $O(1)$
\begin{algorithmic}[1]
\Require $1\leq i,j\leq |V|$ and $t\geq \mathit{currentTime}$ and $w>0$.
\Ensure $G_t$ is up-to-date.
\State $\mathit{currentTime}\gets t$;
\State $z_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.z$;
\State $t_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.t$;
\State $s_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.s$;
\State $z_1\gets z_1+\mathit{currentTime}-t_1$;
\If {$z_1>0$} \State $z_1\gets 0$;\EndIf
\State $z_1\gets z_1-\eta\cdot w$;
\State $z_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.z$;
\State $t_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.t$;
\State $z_2\gets z_2+\mathit{currentTime}-t_2$;
\State $s_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.s$;
\If {$z_2>0$} \State $z_2\gets 0$;\EndIf
\State $z_2\gets z_2-\eta\cdot w$;
\State append ($j$, $t$, $w$, $s_2+w$, $z_2$) to \textit{edgeList}[$i$];
\State append ($i$, $t$, $w$, $s_1+w$, $z_1$) to \textit{edgeList}[$j$];
\If{ binary search on \textit{edgeListReduced}[$i$] did not find $j$}
\State insert $j$ to \textit{edgeListReduced}[$i$] in appropriate position;
\State binary search $i$ on \textit{edgeListReduced}[$j$];
\State insert $i$ to \textit{edgeListReduced}[$j$] in appropriate position;
\EndIf
\State $p\gets \Call{FindRoot}{i}$;
\State $q\gets \Call{FindRoot}{j}$;
\If{$p\neq q$}
\State \Call{Merge}{$p$,$q$}
\EndIf
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htbp]
\caption{\textsc{Find Root}($i$)}\label{alg-find}
\textbf{Time Complexity:} $O(\log |V|)$\\
\textbf{Additional Space:} $O(1)$
\begin{algorithmic}[1]
\Require $1\leq i\leq |V|$
\Ensure return the root of the union tree containing $v_i$.
\State $\mathit{currentTime}\gets t$;
\If {$\mathit{father}[i].u>0$}
	\State \Return{\Call{FindRoot}{$\mathit{father}[i].u$}};
\Else
	\State \Return{$i$};
\EndIf
\end{algorithmic}
\end{algorithm}

@@

\subsection{Computing Metrics}
% TODO
\subsection{Optimization and Parallelizability}
% TODO
\section{Experiment}
\subsection{Data Generation} 
Our data generation is base on real social network provided by \cite{database1,
database2}.  Because the real dynamic graph is hard to obtain, we only use
these graph as a final state of the dynamic graph, which means the evolving
graph generated will satisfy all the property that we discussed in section 2.
Then we artificially add a timeline into that graph to make it become an
evolving graph.

Another way to generate data is to use the method given by
\cite{kumar2010structure}.  The author in that paper propose a method to
simulate the real timeline of a social network, which has same component
properties and very good similarity with the actual network.  Unfortunately,
because of the time limitation, we have not implemented this brilliant
technique.

\subsection{Program}
We have implemented two of the algorithms using C++ language.
The program maintains a dynamic graph which contains $n$ vertices and does not have
any edge at the beginning.
It support the following operations:
\begin{itemize}
	\item \textsc{Add Edge$(u,v,w)$:} increase the time counter by one, and then add an edge
		of weight $w$ between vertex $u$ and vertex $v$.
	\item \textsc{Query Degree$(t,u)$:} ask for the degree of $u$ at time $t$.
	\item \textsc{Query Activity$(t,u)$:} ask for the activity (as we defined) of $u$ at time $t$.
	\item \textsc{Query Connectivity$(t,u,v)$:} ask that whether vertices $u,v$ is connected or not
		$u$ at time $t$.
\end{itemize}
Note that the time counter is initially zero.
Each operation cost $O(\log n)$ time in our implemented algorithm.

@@

\begin{center} \begin{tikzpicture}[y=.3cm, x=.3cm, font=\sffamily]
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (31, 0);
	\draw (0,0) -- coordinate (y axis mid) (0, 19);
	%ticks
	\foreach \x in {2,4,...,30}
		\draw (\x,1pt) -- (\x,-3pt) node[anchor=north] {\scriptsize \x0};
	\foreach \y in {2,4,...,18}
		\draw (1pt,\y) -- (-3pt,\y) node[anchor=east] {\scriptsize \y};
	%labels      
	\node[below=0.5cm] at (x axis mid) {\scriptsize Number of Querys ($\times10^3$)};
	\node[rotate=90, left=0.8cm] at (0, 12) {\scriptsize Running Time (sec.)};
	%plots
	\draw[color=red!70!black] plot[mark=*, mark options={fill=white}] file {exp-n2w};
	\draw[color=green!50!black] plot[mark=square*, mark options={fill=white}] file {exp-n10w};
	\draw[color=blue!80!black] plot[mark=triangle*, mark options={fill=white}] file {exp-n50w};
	%legend
	\begin{scope}[shift={(2,13)}]
	\draw[color=red!70!black] (0,0) -- plot[mark=*, mark options={fill=white}] (1,0) -- (2,0)
		node[right]{\scriptsize number of nodes=20,000};
	\draw[color=green!50!black] (0,1) -- plot[mark=square*, mark options={fill=white}] (1,1) -- (2,1)
		node[right]{\scriptsize number of nodes=100,000};
	\draw[color=blue!90!black] (0,2) -- plot[mark=triangle*, mark options={fill=white}] (1,2) -- (2,2)
		node[right]{\scriptsize number of nodes=500,000};
	\end{scope}
\end{tikzpicture} \end{center}

We can see that the running time grows as linear as number of operations,
	but grows much slower when number of vertices grows (as $\log$).

>>>>

\definition $a_{\mu,\eta}(G_t,v)$ is the \textbf{evolving activity} for node $v$, if
\begin{equation}
a(G_t,v)=\begin{cases}
0& (z\geq 0)\\
\mu^{1/z} & (z<0)
\end{cases},
\end{equation}
where $\mu>1$ is the \textbf{decay rate}, and $z$ is a status variable that can
be determined by the following routine:
\begin{itemize}
\item $z=0$ initially, and $z$ can never exceed $0$.
\item $z$ decrease by $\left(\eta\sum_i w_i\right)$, when new edges with weight
$w_i$'s are created by $v$, as discussed in \eqref{eq-sym}, Here $\eta>0$ is the
\textbf{activity delay}.
\item $z$ increase consistently by time, at the same rate of $t$.
\end{itemize}

@@

Since $\mu>1$, and $z\in\mathbb{R}$, we have $0<a(G_t,v)<1$.
\end{proof}

@@

\section{Data structure and Algorithms}
\subsection{Maintaining the graph}
We formally define the abstract data type for the pure growing model, and
create algorithmic functions to the data structure.

A evolving network $G_t(V,E_t)$ is stored in following contents:
\begin{itemize}
\item \textit{currentTime}, a non-negative real number.
\item $\mu$, a real number in range $(1,+\infty)$.
\item $\eta$, a non-negative real number.
\item \textit{edgeList}[$1\dots|V|$], an array that every of
whose entry points to a sorted array again, consisting the description of the
in edges of $v_i$ in the graph, each with elements:
\begin{itemize}
\item $u$, the other end point of the edge.
\item $t$, the time when this edge is added. (the array is sorted by $t$)
\item $w$, the weight of the edge, can be reduced if the network is
unweighted.
\item $s$, the sum of all previous $w$, including this one.
\item $z$, the status variable for calculating the evolving activity of node
$v_i$ at time $t$ after updating this edge.
\end{itemize}

@@

returns whether $u$ and $v$ is connected in the network at time $t$,
defined in
Algorithm \ref{alg-con}.
\item \textsc{Query Activity}($t,v$)

returns the evolving activity $a_{\mu,\eta}(G_t,v)$, defined in
Algorithm \ref{alg-act}.
\item \textsc{Query Centrality}($t$, $\alpha$, $\beta$, $\mathbf{x}_0$, $m$)

returns the active centrality vector $\mathbf{x}(G_t)$. Algorithm
\ref{alg-cen} use the iteration method start with $\mathbf{x}_0$ and for
$m$ times.
\item \textsc{Modify Mu}($\mu'$)

change the value of $\mu$ which defines the evolving activity, in case the
current $\mu$ is inappropriate, defined in algorithm \ref{alg-mu}.
\item \textsc{Modify Eta}($\eta'$)

change the value of $\eta$ which defines the evolving activity, in case the
current $\eta$ is inappropriate, defined in algorithm \ref{alg-eta}.
\end{itemize}

\begin{algorithm}[htbp]
\caption{\textsc{Add Edge}($v_i,v_j,t,w$)}
\label{alg-add}
\textbf{Time Complexity:} $O(\log |V|)$\\
\textbf{Additional Space:} $O(1)$
\begin{algorithmic}[1]
\Require $1\leq i,j\leq |V|$ and $t\geq \mathit{currentTime}$ and $w>0$.
\Ensure $G_t$ is up-to-date.
\State $\mathit{currentTime}\gets t$;
\State $z_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.z$;
\State $t_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.t$;
\State $s_1\gets\mathit{edgeList}[i].\mathrm{lastElement}.s$;
\State $z_1\gets z_1+\mathit{currentTime}-t_1$;
\If {$z_1>0$} \State $z_1\gets 0$;\EndIf
\State $z_1\gets z_1-\eta\cdot w$;
\State $z_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.z$;
\State $t_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.t$;
\State $z_2\gets z_2+\mathit{currentTime}-t_2$;
\State $s_2\gets\mathit{edgeList}[j].\mathrm{lastElement}.s$;
\If {$z_2>0$} \State $z_2\gets 0$;\EndIf
\State $z_2\gets z_2-\eta\cdot w$;
\State append ($j$, $t$, $w$, $s_2+w$, $z_2$) to \textit{edgeList}[$i$];
\State append ($i$, $t$, $w$, $s_1+w$, $z_1$) to \textit{edgeList}[$j$];
\If{ binary search on \textit{edgeListReduced}[$i$] did not find $j$}
\State insert $j$ to \textit{edgeListReduced}[$i$] in appropriate position;
\State binary search $i$ on \textit{edgeListReduced}[$j$];
\State insert $i$ to \textit{edgeListReduced}[$j$] in appropriate position;
\EndIf
\State $p\gets \Call{FindRoot}{i,\mathit{currentTime}}$;
\State $q\gets \Call{FindRoot}{j,\mathit{currentTime}}$;
\If{$p\neq q$}
\State \Call{Merge}{$p$,$q$}
\EndIf
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htbp]
\caption{\textsc{Find Root}($i$, $t$)}\label{alg-find}
\textbf{Time Complexity:} $O(\log |V|)$\\
\textbf{Additional Space:} $O(1)$
\begin{algorithmic}[1]
\Require $1\leq i\leq |V|$
\Ensure return the root of the union tree containing $v_i$ at time $t$.
\State $\mathit{currentTime}\gets t$;
\If {$\mathit{father}[i].u>0$ and $\mathit{father}[i].t\leq t$}
	\State \Return{\Call{FindRoot}{$\mathit{father}[i].u$}};
\Else
	\State \Return{$i$};
\EndIf
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[htbp]
\caption{\textsc{Query Connectivity}($t$,$p$,$q$)}
\label{alg-con}
\textbf{Time Complexity:} $O(\log |V|)$\\
\textbf{Additional Space:} $O(1)$
\begin{algorithmic}[1]
\Require $1\leq p,q \leq |V|$.
\Ensure return whether $p$ and $q$ are connected at time $t$.
\State \Return{$\Call{Find Root}{p,t}=\Call{Find Root}{q,t}$};
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htbp]
\caption{\textsc{Query Degree}($t$,$v_i$)}
\label{alg-deg}
\textbf{Time Complexity:} $O(\log |V|)$\\
\textbf{Additional Space:} $O(1)$
\begin{algorithmic}[1]
\Require $1\leq i\leq |V|$.
\Ensure return the degree of $v_i$ at time $t$.
\State $p\gets\text{binary search }\mathit{edgeList}[i]\text{ for
the latest edge before }t$;
\If{ $p$ not found}
\State\Return{$0$}
\Else
\State \Return{$p.s$};
\EndIf
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htbp]
\caption{\textsc{Query Activity}($t$,$v_i$)}
\label{alg-act}
\textbf{Time Complexity:} $O(\log |V|)$\\
\textbf{Additional Space:} $O(1)$
\begin{algorithmic}[1]
\Require $1\leq i\leq |V|$.
\Ensure return the evolving activity $a_{\mu,\eta}(G_t,v_i)$.
\State $p\gets\text{binary search }\mathit{edgeList}[i]\text{ for
the latest edge before }t$;
\If{ $p$ not found}
\State\Return{$0$}
\Else
\State $z=\min(p.z+t-p.t,0)$;
\If{$z=0$}
\State \Return{$0$};
\Else
\State \Return{$\mu^{1/z}$};
\EndIf
\EndIf
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htbp]
\caption{\textsc{Modify Mu}($\mu'$)}
\label{alg-mu}
\textbf{Time Complexity:} $O(1)$\\
\textbf{Additional Space:} $O(1)$
\begin{algorithmic}[1]
\State $\mu\gets\mu'$;
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htbp]
\caption{\textsc{Modify Eta}($\eta'$)}
\label{alg-eta}
\textbf{Time Complexity:} $O(1)$\\
\textbf{Additional Space:} $O(1)$
\begin{algorithmic}[1]
\State $\eta\gets\eta'$;\State\Comment{Imcompatible with historical queries.
The alternative is to rebuild the whole graph.}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htbp]
\caption{\textsc{Query Centrality}($t$,$\alpha$,$\beta$,
$\mathbf{x}_0$,$m$)}
\label{alg-cen}
\textbf{Time Complexity:} $O(mK|V|)$, $K$ is the sparsity constant.\\
\textbf{Additional Space:} $O(|V|)$
\begin{algorithmic}[1]
\Require $0<\alpha<\lambda_1(A^\bullet(G_t))^{-1}$ and $\beta>0$ and $m>0$. Here
$\lambda_1(A)$ is the maximal possitive eigenvalue of matrix $A$.
\Ensure return the active centrality $\mathbf{x}(G_t)$.
\For{$i\gets 1\dots|V|$}
\State $a[i]\gets \Call{Query Activity}{t,v_i}$;
\State $\mathbf{x}[i]\gets \mathbf{x}_0[i]$;
\EndFor
\For{$\mathit{count}\gets 1\dots m$}
\For{$i\gets 1\dots|V|$}
\State $\mathbf{x'}[i]\gets \beta$;
\EndFor
\For{$i\gets 1\dots|V|$}
\ForAll{$j\in\mathit{edgeListReduced}[i]$}
\State $\mathbf{x'}[i]\gets \mathbf{x'}[i]+\alpha\cdot a[j]\cdot\mathbf{x}[j]$
\EndFor\EndFor
\For{$i\gets 1\dots|V|$}
\State $\mathbf{x}[i]\gets \mathbf{x'}[i]$;
\EndFor
\EndFor
\State\Return{$\mathbf{x}$};
\end{algorithmic}
\end{algorithm}

\subsection{Optimization and Parallelizability}

In most cases, we only need to query the latest properties of the graph, i.e.,
$t=\mathit{currentTime}$. And we observe that in Algorithm \ref{alg-con},
\ref{alg-deg} and \ref{alg-act}, the binary search can be replace by simply find
the last entry of the argument sequence. So those three algorithms can be done
in $O(1)$ time if we do not perform historical queries.

Algorithm \ref{alg-cen} actually repeat the computation of the multiplication
of a sparse matrix to a vector, which is highly parallelizable[?]. So in a large
scale distributed system, we can perform very efficient analysis to an evolving
graph, by our model.
\section{Experiment}
\subsection{Data Generation} 
Our data generation is base on real social networks provided by \cite{database1,
database2}. Unfortunatly, those databases do not contain evolving history
recorded data, so we simulate the evolving history by use the
authentic graph as the final state of the tested dynamic graph.
We give a random permutation to the edges, in a way such that there will always
be just one component -- known as the giant component. Because the final state
is from real world data, it is ensured 
the evolving graph generated by our model will satisfy all the property that we
discussed in section 2.2. In addition, Kumar introduced another way to generate
artificial data of evolving social networks\cite{kumar2010structure}.

\subsection{Program}
We have implemented the algorithms using C++ language without compiler
optimization, and all execution is done under a typical PC with Intel Core i3.

The program maintains a dynamic graph which contains $n$ vertices and does not
have any edge at the beginning.
It support the following operations:
\begin{itemize}
	\item \textsc{Add Edge$(u,v,w)$:} increase the time counter by one, and then add an edge
		of weight $w$ between vertex $u$ and vertex $v$.
	\item \textsc{Query Degree$(t,u)$:} ask for the degree of $u$ at time $t$.
	\item \textsc{Query Activity$(t,u)$:} ask for the activity (as we defined) of $u$ at time $t$.
	\item \textsc{Query Connectivity$(t,u,v)$:} ask that whether vertices $u,v$ is connected or not
		$u$ at time $t$.
\end{itemize}
Note that the time counter is initially zero.
Each operation cost $O(\log n)$ time in our implemented algorithm.

@@

\begin{center} \begin{tikzpicture}[y=.3cm, x=.3cm, font=\sffamily]
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (31, 0);
	\draw (0,0) -- coordinate (y axis mid) (0, 19);
	%ticks
	\foreach \x in {2,4,...,30}
		\draw (\x,1pt) -- (\x,-3pt) node[anchor=north] {\scriptsize \x};
	\foreach \y in {2,4,...,18}
		\draw (1pt,\y) -- (-3pt,\y) node[anchor=east] {\scriptsize \y};
	%labels      
	\node[below=0.5cm] at (x axis mid) {\scriptsize Number of Queries ($\times10^5$)};
	\node[rotate=90, left=0.8cm] at (0, 12) {\scriptsize Running Time (sec.)};
	%plots
	\draw[color=red!70!black] plot[mark=*, mark options={fill=white}] file {exp-n2w};
	\draw[color=green!50!black] plot[mark=square*, mark options={fill=white}] file {exp-n10w};
	\draw[color=blue!80!black] plot[mark=triangle*, mark options={fill=white}] file {exp-n50w};
	%legend
	\begin{scope}[shift={(2,13)}]
	\draw[color=red!70!black] (0,0) -- plot[mark=*, mark options={fill=white}] (1,0) -- (2,0)
		node[right]{\scriptsize number of nodes=20,000};
	\draw[color=green!50!black] (0,1) -- plot[mark=square*, mark options={fill=white}] (1,1) -- (2,1)
		node[right]{\scriptsize number of nodes=100,000};
	\draw[color=blue!90!black] (0,2) -- plot[mark=triangle*, mark options={fill=white}] (1,2) -- (2,2)
		node[right]{\scriptsize number of nodes=500,000};
	\end{scope}
\end{tikzpicture} \end{center}

We can verify that our algorithm model works efficiently with those four kinds
of basic operations.
